---
title: "Plots for our `surveyCV` *Stat* paper based on SDSS presentation"
author: "Cole Guerin, Thomas McMahon, Jerzy Wieczorek"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plots-for-Stat-paper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Writing cleaner code for **only** the simulations needed for our submission to *Stat* (the special issue for SDSS 2021 talks).

Currently based on the sims in `plot_generation.R` -- but instead of having similar functions repeatedly defined separately in that file, use our actual functions from the package such as `cv.svy()`.

* TODO: Rewrite `Sims.Weight.plot` to save the sim results, and make the plots separately

```{r Packages, message=FALSE}
library(survey)
library(ggplot2)
library(splines)
library(magrittr)
```
```{r LoadSurveyCV}
library(surveyCV)
```

## Generate the artificial population

```{r ArtifPop}
set.seed(47)

x1 = stats::runif(1:500, min = 26, max = 38)
y1 = (x1-29)^3 - 13*(x1-29)^2 + 0*(x1-29) + 900

x2 = stats::runif(1:500, min = 38, max = 50)
y2 = (x2-36)^3 - 10*(x2-36)^2 + 2*(x2-36) + 600

z1 = jitter(y1, 15000)
z2 = jitter(y2, 15000)

ds1 <- data.frame(Response = z1, Predictor = x1)
ds2 <- data.frame(Response = z2, Predictor = x2)
ds12 <- rbind(ds1, ds2)

b12 <- data.frame(ID = c(1:1000))
splinepop.df <- cbind(b12, ds12)
splinepop.df <- splinepop.df %>%
  dplyr::arrange(Predictor) %>%
  dplyr::mutate(Stratum = dplyr::row_number(),
                Cluster = dplyr::row_number()) %>% 
  dplyr::mutate(Stratum = cut(Stratum, 5, 1:5),
                Cluster = cut(Cluster, 100, 1:100)) %>%
  # full pop has 1000 units; 200 units/stratum; 100 clusters of 10 units each
  dplyr::mutate(fpcSRS = 1000,
                fpcStratum = 1000/5,
                fpcCluster = 100) %>%
  dplyr::arrange(ID)

lm_quad <- stats::lm(Response ~ Predictor + I(Predictor^2),
                     data = splinepop.df)
splinepop.df$samp_prob_quad <-
  (1/(abs(lm_quad$residuals))) / sum(1/(abs(lm_quad$residuals)))
splinepop.df$samp_wt_quad <- 1/splinepop.df$samp_prob_quad

quad.sample.graph <- ggplot(splinepop.df,
  aes(x = Predictor, y = Response)) +
  geom_point(aes(size = samp_prob_quad), alpha = 0.2) +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), se = FALSE)
```

```{r QuadWeightsPlot, fig.width=7, fig.height=7}
quad.sample.graph
```


## Sims: Use of sampling weights (in training model-fits vs in test loss-estimates)


```{r WeightsSimsFunction}
Sims.Weight.plot <- function(n, loops, plot, weights) {
  stopifnot(all.equal(sum(1/splinepop.df[[weights]]), 1))

  # In all the following cases,
  # we are only checking the effect of using sampling **weights**
  # during model-fitting on training sets, and testing on test sets;
  # we did not use cluster or stratified sampling,
  # so all CV folds will (sensibly) be SRS (regardless of samp weights)
  
  # Use weighted models, and calculate MSEs using weighted design
  AllW <- data.frame(df = c(), MSE = c())
  # Use SRS models, and calculate MSEs using SRS design
  NoW <- data.frame(df = c(), MSE = c())
  # Use weighted models, and calculate MSEs using SRS design
  ModW <- data.frame(df = c(), MSE = c())
  # Use SRS models, and calculate MSEs using weighted design
  MSEW <- data.frame(df = c(), MSE = c())
  
  for (i in 1:loops) {
    # Take a sample of size n, using the sampling probabilities instead of SRS
    # (using 1/samp_wt as the samp_prob)
    in.sample <- sampling::UPtille(n / splinepop.df[[weights]])
    splinesamp.df <- splinepop.df[in.sample > 0, ]
    modelsToFit <- c("Response~ns(Predictor, df=1)", "Response~ns(Predictor, df=2)",
                     "Response~ns(Predictor, df=3)", "Response~ns(Predictor, df=4)",
                     "Response~ns(Predictor, df=5)", "Response~ns(Predictor, df=6)")

    AllWdat <- cv.svy(splinesamp.df, modelsToFit,
                      nfolds = 5, weightsID = weights) %>% as.vector()
    NoWdat <- cv.svy(splinesamp.df, modelsToFit,
                     nfolds = 5) %>% as.vector()
    ModWdat <- cv.svy(splinesamp.df, modelsToFit,
                      nfolds = 5, weightsID = weights,
                      useSvyForLoss = FALSE) %>% as.vector()
    MSEWdat <- cv.svy(splinesamp.df, modelsToFit,
                      nfolds = 5, weightsID = weights,
                      useSvyForFits = FALSE) %>% as.vector()

    # compiling one data frame
    AllW2 <- data.frame(df = 1:6, MSE = AllWdat, sample = rep(i, 6))
    AllW <- rbind(AllW, AllW2)
    NoW2 <- data.frame(df = 1:6, MSE = NoWdat, sample = rep(i, 6))
    NoW <- rbind(NoW, NoW2)
    ModW2 <- data.frame(df = 1:6, MSE = ModWdat, sample = rep(i, 6))
    ModW <- rbind(ModW, ModW2)
    MSEW2 <- data.frame(df = 1:6, MSE = MSEWdat, sample = rep(i, 6))
    MSEW <- rbind(MSEW, MSEW2)
  }
  # Making the degrees of freedom variable a factor variable
  AllW$df <- as.factor(AllW$df)
  NoW$df <- as.factor(NoW$df)
  MSEW$df <- as.factor(MSEW$df)
  ModW$df <- as.factor(ModW$df)
  # Find the y-range of MSEs
  ymin <- min(min(AllW$MSE), min(NoW$MSE), min(MSEW$MSE), min(ModW$MSE))
  ymax <- max(max(AllW$MSE), max(NoW$MSE), max(MSEW$MSE), max(ModW$MSE))
  # Making a ggplot object for the MSEs collected when using
  # SRS folds, SRS models, and SRS error calculations
  p1 <- ggplot(data = AllW, mapping = aes(x = df, y = MSE)) +
    ggtitle("Weights for both") +
    ylim(ymin, ymax)
  # Making a ggplot object for the MSEs collected when using
  # SRS folds, Clus models, and SRS error calculations
  p2 <- ggplot(data = NoW, mapping = aes(x = df, y = MSE)) +
    ggtitle("No Weights") +
    ylim(ymin, ymax)
  # Making a ggplot object for the MSEs collected when using
  # Clus folds, Clus models, and Clus error calculations
  p3 <- ggplot(data = ModW, mapping = aes(x = df, y = MSE)) +
    ggtitle("Weights when modeling") +
    ylim(ymin, ymax)
  # Making a ggplot object for the MSEs collected when using
  # Clus folds, SRS models, and Clus error calculations
  p4 <- ggplot(data = MSEW, mapping = aes(x = df, y = MSE)) +
    ggtitle("Weights when MSE gen") +
    ylim(ymin, ymax)

  # Either turning our ggplot objects into boxplots or spaghetti plots
  # (as objects still)
  if (plot == "box") {
    plot1 <- p1 + geom_boxplot()
    plot2 <- p2 + geom_boxplot()
    plot3 <- p3 + geom_boxplot()
    plot4 <- p4 + geom_boxplot()
  } else if (plot == "line") {
    plot1 <- p1 + geom_line(aes(group = sample, colour = sample)) +
      theme(legend.position = "none")
    plot2 <- p2 + geom_line(aes(group = sample, colour = sample)) +
      theme(legend.position = "none")
    plot3 <- p3 + geom_line(aes(group = sample, colour = sample)) +
      theme(legend.position = "none")
    plot4 <- p4 + geom_line(aes(group = sample, colour = sample)) +
      theme(legend.position = "none")
  }
  # Making a grid display of the four plot objects above
  gridExtra::grid.arrange(plot1,plot2,plot3,plot4, ncol = 2,
               top = paste0("Simulated Spline Data (Sample Size = ", n,
                            ", Loops = ", loops, ", Weights = ", weights, ")"))
}
```

```{r WeightsSimsResultsAndPlots, fig.width=7, fig.height=7}
Main.weights.plot <- Sims.Weight.plot(200, 10, "box", "samp_wt_quad")
```




## Sims: Use of surveyCV folds with SRS, clustered, or stratified samples

```{r FoldsSimsSetup}
n <- 100
loops <- 100
```

```{r FoldsSimsResults, eval = FALSE}
# Use SRS samples and make SRS folds
srssrsds <- data.frame(df = c(), MSE = c())
# Use SRS samples and use dAIC
srsdAICds <- data.frame(df = c(), dAIC = c())

# Use Cluster samples and make Cluster folds
clusclusds <- data.frame(df = c(), MSE = c())
# Use Cluster samples and make SRS folds
clussrsds <- data.frame(df = c(), MSE = c())
# Use Cluster samples and use dAIC
clusdAICds <- data.frame(df = c(), dAIC = c())

# Use Strat samples and make Strat folds
stratstratds <- data.frame(df = c(), MSE = c())
# Use Strat samples and make SRS folds
stratsrsds <- data.frame(df = c(), MSE = c())
# Use Strat samples and use dAIC
stratdAICds <- data.frame(df = c(), dAIC = c())

modelsToFit <- c("Response~ns(Predictor, df=1)", "Response~ns(Predictor, df=2)",
                 "Response~ns(Predictor, df=3)", "Response~ns(Predictor, df=4)",
                 "Response~ns(Predictor, df=5)", "Response~ns(Predictor, df=6)")

# Making as many simple random samples as we specify for 'loops'
for (i in 1:loops) {
  # Take a SRS
  set.seed(i)
  sim.srs <- dplyr::sample_n(splinepop.df, n)
  
  # Using our SRS function on SRS samples
  srs.CV.out <- cv.svy(sim.srs, modelsToFit,
                       nfolds = 5, fpcID = "fpcSRS") %>% as.vector()
  srssrsds2 <- data.frame(df = 1:6, MSE = srs.CV.out)
  srssrsds <- rbind(srssrsds, srssrsds2)

  # Using dAIC on SRS samples
  srs.des <- svydesign(ids = ~1, fpc = ~fpcSRS, data = sim.srs)
  srs.dAIC.out <- sapply(1:6,
    function(ii) AIC(svyglm(formula = formula(modelsToFit[ii]),
                            design = srs.des))["AIC"])
  srsdAICds2 <- data.frame(df = 1:6, dAIC = srs.dAIC.out)
  srsdAICds <- rbind(srsdAICds, srsdAICds2)
}

# Making as many cluster samples as we specify for 'loops'
for (i in 1:loops) {
  # Take a Cluster sample
  set.seed(i)
  c <- unique(splinepop.df[["Cluster"]])
  sim.clus <- splinepop.df[splinepop.df[["Cluster"]] %in% sample(c, n/10),]
  
  # Using our Cluster function on Cluster samples
  clus.CV.out <- cv.svy(sim.clus, modelsToFit,
                        clusterID = "Cluster", nfolds = 5, fpcID = "fpcCluster") %>% as.vector()
  clusclusds2 <- data.frame(df = 1:6, MSE = clus.CV.out)
  clusclusds <- rbind(clusclusds, clusclusds2)

  # Using our SRS function on Cluster samples
  srs.CV.out <- cv.svy(sim.clus, modelsToFit,
                       nfolds = 5, fpcID = "fpcSRS") %>% as.vector()
  clussrsds2 <- data.frame(df = 1:6, MSE = srs.CV.out)
  clussrsds <- rbind(clussrsds, clussrsds2)
  
  # Using dAIC on Cluster samples
  clus.des <- svydesign(ids = ~Cluster, fpc = ~fpcCluster, data = sim.clus)
  clus.dAIC.out <- sapply(1:6,
    function(ii) AIC(svyglm(modelsToFit[ii], clus.des))["AIC"])
  clusdAICds2 <- data.frame(df = 1:6, dAIC = clus.dAIC.out)
  clusdAICds <- rbind(clusdAICds, clusdAICds2)
}

# Making as many stratified samples as we specify for 'loops'
for (i in 1:loops) {
  # Take a Strat sample
  set.seed(i)
  s <- survey::stratsample(splinepop.df$Stratum,
         c("1" = n/5, "2" = n/5, "3" = n/5, "4" = n/5, "5" = n/5))
  sim.strat <- splinepop.df[s,]
  
  # Using our Strat function on Strat samples
  strat.CV.out <- cv.svy(sim.strat, modelsToFit,
                         strataID = "Stratum", nfolds = 5, fpcID = "fpcStratum") %>% as.vector()
  stratstratds2 <- data.frame(df = 1:6, MSE = strat.CV.out)
  stratstratds <- rbind(stratstratds, stratstratds2)
  
  # Using our SRS function on Strat samples
  srs.CV.out <- cv.svy(sim.strat, modelsToFit,
                       nfolds = 5, fpcID = "fpcSRS") %>% as.vector()
  stratsrsds2 <- data.frame(df = 1:6, MSE = srs.CV.out)
  stratsrsds <- rbind(stratsrsds, stratsrsds2)
  
  # Using dAIC on Strat samples
  strat.des <- svydesign(ids = ~1, strata = ~Stratum, fpc = ~fpcStratum, data = sim.strat)
  strat.dAIC.out <- sapply(1:6,
    function(ii) AIC(svyglm(modelsToFit[ii], strat.des))["AIC"])
  stratdAICds2 <- data.frame(df = 1:6, dAIC = strat.dAIC.out)
  stratdAICds <- rbind(stratdAICds, stratdAICds2)
}

# Making the degrees of freedom variable a factor variable
srssrsds$df <- as.factor(srssrsds$df)
srsdAICds$df <- as.factor(srsdAICds$df)
clusclusds$df <- as.factor(clusclusds$df)
clussrsds$df <- as.factor(clussrsds$df)
clusdAICds$df <- as.factor(clusdAICds$df)
stratstratds$df <- as.factor(stratstratds$df)
stratsrsds$df <- as.factor(stratsrsds$df)
stratdAICds$df <- as.factor(stratdAICds$df)

usethis::use_data(srssrsds, srsdAICds,
                  clusclusds, clussrsds, clusdAICds,
                  stratstratds, stratsrsds, stratdAICds,
                  internal = FALSE)
# for now setting internal=FALSE
# so that we don't overwrite the sysdata.Rda for intro.Rmd;
# but eventually when we're ready to replace that old vignette,
# we can return to internal=TRUE here
# if that's indeed better for R packages
```


```{r FoldsSimsPlots, fig.width=7, fig.height=7}
data("srssrsds",
     "clusclusds", "clussrsds",
     "stratstratds", "stratsrsds",
     "srsdAICds", "clusdAICds", "stratdAICds")

# Find the y-ranges of MSEs and dAICs
yminMSE <- min(min(srssrsds$MSE),
               min(clusclusds$MSE), min(clussrsds$MSE),
               min(stratstratds$MSE), min(stratsrsds$MSE))
ymaxMSE <- max(max(srssrsds$MSE),
               max(clusclusds$MSE), max(clussrsds$MSE),
               max(stratstratds$MSE), max(stratsrsds$MSE))
yminAIC <- min(min(srsdAICds$dAIC), min(clusdAICds$dAIC),
               min(stratdAICds$dAIC))
ymaxAIC <- max(max(srsdAICds$dAIC), max(clusdAICds$dAIC),
               max(stratdAICds$dAIC))

# Making ggplot objects for the MSEs and AICs
plot.srssrs <- ggplot(data = srssrsds, mapping = aes(x = df, y = MSE/1e4)) +
  geom_boxplot() +
  ggtitle("SRS folds,\nSRS sample") +
  scale_y_log10(limits = c(yminMSE, ymaxMSE)/1e4)
plot.blank <- rectGrob(width = 0, height = 0) # empty spot here -- no corresponding sim
plot.clusclus <- ggplot(data = clusclusds, mapping = aes(x = df, y = MSE/1e4)) +
  geom_boxplot() +
  ggtitle("Cluster folds,\nCluster sample") +
  scale_y_log10(limits = c(yminMSE, ymaxMSE)/1e4)
plot.clussrs <- ggplot(data = clussrsds, mapping = aes(x = df, y = MSE/1e4)) +
  geom_boxplot() +
  ggtitle("SRS folds,\nCluster sample") +
  scale_y_log10(limits = c(yminMSE, ymaxMSE)/1e4)
plot.stratstrat <- ggplot(data = stratstratds, mapping = aes(x = df, y = MSE/1e4)) +
  geom_boxplot() +
  ggtitle("Strat folds,\nStrat sample") +
  scale_y_log10(limits = c(yminMSE, ymaxMSE)/1e4)
plot.stratsrs <- ggplot(data = stratsrsds, mapping = aes(x = df, y = MSE/1e4)) +
  geom_boxplot() +
  ggtitle("SRS folds,\nStrat sample") +
  scale_y_log10(limits = c(yminMSE, ymaxMSE)/1e4)

plot.srsdAIC <- ggplot(data = srsdAICds, mapping = aes(x = df, y = dAIC/1e6)) +
  geom_boxplot() +
  ggtitle("dAIC,\nSRS sample") +
  scale_y_log10(limits = c(yminAIC, ymaxAIC)/1e6)
plot.clusdAIC <- ggplot(data = clusdAICds, mapping = aes(x = df, y = dAIC/1e6)) +
  geom_boxplot() +
  ggtitle("dAIC,\nCluster sample") +
  scale_y_log10(limits = c(yminAIC, ymaxAIC)/1e6)
plot.stratdAIC <- ggplot(data = stratdAICds, mapping = aes(x = df, y = dAIC/1e6)) +
  geom_boxplot() +
  ggtitle("dAIC,\nStrat sample") +
  scale_y_log10(limits = c(yminAIC, ymaxAIC)/1e6)


# Making a grid display of the plot objects above
gridExtra::grid.arrange(plot.srsdAIC, plot.srssrs, plot.blank,
                        plot.clusdAIC, plot.clussrs, plot.clusclus,
                        plot.stratdAIC, plot.stratsrs, plot.stratstrat,
             ncol = 3,
             top = paste0("Simulated Spline Data (Sample Sizes = ", n,
                          ", Clusters = ", n/10,
                          " or Strata = 5, Loops = ", loops, ", Folds = 5)"))
```

