---
title: "Plots for our `surveyCV` *Stat* paper based on SDSS presentation"
author: "Cole Guerin, Thomas McMahon, Jerzy Wieczorek"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plots-for-Stat-paper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Writing cleaner code for **only** the simulations needed for our submission to *Stat* (the special issue for SDSS 2021 talks).

Currently based on the sims in `plot_generation.R` -- but instead of having similar functions repeatedly defined separately in that file, use our actual functions from the package such as `cv.svy()`.

* TODO: Fix the use of ` %>% print()`

```{r Packages, message=FALSE}
library(survey)
library(ggplot2)
library(splines)
library(magrittr)
```
```{r LoadSurveyCV}
library(surveyCV)
```

## Generate the artificial population

```{r ArtifPop}
set.seed(47)
x1 = stats::runif(1:500, min = 26, max = 38)
y1 = (x1-29)^3 - 13*(x1-29)^2 + 0*(x1-29) + 900

set.seed(47)
x2 = stats::runif(1:500, min = 38, max = 50)
y2 = (x2-36)^3 - 10*(x2-36)^2 + 2*(x2-36) + 600

set.seed(47)
# TODO: remove these 1st few duplicate lines eventually --
# but would mess up the random seed, so don't remove yet
# until we've tested and can replicate old plots first...
# THEN when we are happy with ability to replicate old result,
# revise so that both sets of sims (eval folds AND eval weights)
# are using the SAME simulated dataset.
z1 = jitter(y1, 15000)
z1 = jitter(y1, 15000)
z2 = jitter(y2, 15000)
z3 = jitter(y1, 15000)
z4 = jitter(y2, 15000)

ds1 <- data.frame(Response = z1, Predictor = x1)
ds2 <- data.frame(Response = z2, Predictor = x2)
ds12 <- rbind(ds1, ds2)

ds3 <- data.frame(Response = z3, Predictor = x1)
ds4 <- data.frame(Response = z4, Predictor = x2)
ds34 <- rbind(ds3, ds4)

b12 <- data.frame(ID = c(1:1000))
spline.df2 <- cbind(b12, ds12)
spline.df2 <- spline.df2 %>%
  dplyr::arrange(Predictor) %>%
  dplyr::mutate(Stratum = dplyr::row_number(),
                Cluster = dplyr::row_number())
spline.df2$Stratum <- cut(spline.df2$Stratum, 5, 1:5)
spline.df2$Cluster <- cut(spline.df2$Cluster, 100, 1:100)
spline.df2 <- spline.df2 %>%
  dplyr::arrange(ID) %>%
  dplyr::select(ID, Response, Predictor, Cluster, Stratum)




# TODO: remove this eventually (once we can replicate old results)
# and just add samp_wt_quad to the original artificial pop above
b34 <- data.frame(ID = c(1:1000))
spline.df3 <- cbind(b34, ds34)
spline.df3 <- spline.df3 %>%
  dplyr::arrange(Predictor) %>%
  dplyr::mutate(Stratum = dplyr::row_number(),
                Cluster = dplyr::row_number())
spline.df3$Stratum <- cut(spline.df3$Stratum, 5, 1:5)
spline.df3$Cluster <- cut(spline.df3$Cluster, 100, 1:100)
spline.df3 <- spline.df3 %>%
  dplyr::arrange(ID) %>%
  dplyr::select(ID, Response, Predictor, Cluster, Stratum)

lm_quad <- stats::lm(Response ~ Predictor + I(Predictor^2),
                     data = spline.df3)
spline.df3$samp_prob_quad <-
  (1/(abs(lm_quad$residuals))) / sum(1/(abs(lm_quad$residuals)))
spline.df3$samp_wt_quad <- 1/spline.df3$samp_prob_quad

quad.sample.graph <- ggplot(spline.df3,
  aes(x = Predictor, y = Response)) +
  geom_point(aes(size = samp_prob_quad), alpha = 0.2) +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), se = FALSE)
```

```{r QuadWeightsPlot, fig.width=7, fig.height=7}
quad.sample.graph
```


## Sims: Use of sampling weights (in training model-fits vs in test loss-estimates)


```{r WeightsSimsFunction}
# TODO: rename spline.df3 to be more clearly
# the sim dataset used for testing effect of weights

# # TEMPORARILY check that using different artif pop DOES give diff results?
# spline.df3$Predictor <- spline.df2$Predictor
# spline.df3$Response <- spline.df2$Response
# # YES it does... Quite substantially!
# # TODO: we ought to run many more than just 10 loops in the function below
# # before we can safely interpret the results.

Sims.Weight.plot <- function(n, loops, plot, weights) {
  stopifnot(all.equal(sum(1/spline.df3[[weights]]), 1))

  # In all the following cases,
  # we are only checking the effect of using sampling **weights**
  # during model-fitting on training sets, and testing on test sets;
  # we did not use cluster or stratified sampling,
  # so all CV folds will (sensibly) be SRS (regardless of samp weights)
  
  # Use weighted models, and calculate MSEs using weighted design
  AllW <- data.frame(df = c(), MSE = c())
  # Use SRS models, and calculate MSEs using SRS design
  NoW <- data.frame(df = c(), MSE = c())
  # Use weighted models, and calculate MSEs using SRS design
  ModW <- data.frame(df = c(), MSE = c())
  # Use SRS models, and calculate MSEs using weighted design
  MSEW <- data.frame(df = c(), MSE = c())
  
  for (i in 1:loops) {
    # Take a sample of size n, using the sampling probabilities instead of SRS
    # (using 1/samp_wt as the samp_prob)
    in.sample <- sampling::UPtille(n / spline.df3[[weights]])
    spline.df3.sample <- spline.df3[in.sample > 0, ]
    modelsToFit <- c("Response~ns(Predictor, df=1)", "Response~ns(Predictor, df=2)",
                     "Response~ns(Predictor, df=3)", "Response~ns(Predictor, df=4)",
                     "Response~ns(Predictor, df=5)", "Response~ns(Predictor, df=6)")

    AllWdat <- cv.svy(spline.df3.sample, modelsToFit,
                      nfolds = 5, weightsID = weights) %>% print()
    NoWdat <- cv.svy(spline.df3.sample, modelsToFit,
                     nfolds = 5) %>% print()
    ModWdat <- cv.svy(spline.df3.sample, modelsToFit,
                      nfolds = 5, weightsID = weights,
                      useSvyForLoss = FALSE) %>% print()
    MSEWdat <- cv.svy(spline.df3.sample, modelsToFit,
                      nfolds = 5, weightsID = weights,
                      useSvyForFits = FALSE) %>% print()

    # compiling one data frame
    AllW2 <- data.frame(df = 1:6, MSE = AllWdat[,1], sample = rep(i, 6))
    AllW <- rbind(AllW, AllW2)
    NoW2 <- data.frame(df = 1:6, MSE = NoWdat[,1], sample = rep(i, 6))
    NoW <- rbind(NoW, NoW2)
    ModW2 <- data.frame(df = 1:6, MSE = ModWdat[,1], sample = rep(i, 6))
    ModW <- rbind(ModW, ModW2)
    MSEW2 <- data.frame(df = 1:6, MSE = MSEWdat[,1], sample = rep(i, 6))
    MSEW <- rbind(MSEW, MSEW2)
  }
  # Making the degrees of freedom variable a factor variable
  AllW$df <- as.factor(AllW$df)
  NoW$df <- as.factor(NoW$df)
  MSEW$df <- as.factor(MSEW$df)
  ModW$df <- as.factor(ModW$df)
  # Find the y-range of MSEs
  ymin <- min(min(AllW$MSE), min(NoW$MSE), min(MSEW$MSE), min(ModW$MSE))
  ymax <- max(max(AllW$MSE), max(NoW$MSE), max(MSEW$MSE), max(ModW$MSE))
  # Making a ggplot object for the MSEs collected when using
  # SRS folds, SRS models, and SRS error calculations
  p1 <- ggplot(data = AllW, mapping = aes(x = df, y = MSE)) +
    ggtitle("Weights for both") +
    ylim(ymin, ymax)
  # Making a ggplot object for the MSEs collected when using
  # SRS folds, Clus models, and SRS error calculations
  p2 <- ggplot(data = NoW, mapping = aes(x = df, y = MSE)) +
    ggtitle("No Weights") +
    ylim(ymin, ymax)
  # Making a ggplot object for the MSEs collected when using
  # Clus folds, Clus models, and Clus error calculations
  p3 <- ggplot(data = ModW, mapping = aes(x = df, y = MSE)) +
    ggtitle("Weights when modeling") +
    ylim(ymin, ymax)
  # Making a ggplot object for the MSEs collected when using
  # Clus folds, SRS models, and Clus error calculations
  p4 <- ggplot(data = MSEW, mapping = aes(x = df, y = MSE)) +
    ggtitle("Weights when MSE gen") +
    ylim(ymin, ymax)

  # Either turning our ggplot objects into boxplots or spaghetti plots
  # (as objects still)
  if (plot == "box") {
    plot1 <- p1 + geom_boxplot()
    plot2 <- p2 + geom_boxplot()
    plot3 <- p3 + geom_boxplot()
    plot4 <- p4 + geom_boxplot()
  } else if (plot == "line") {
    plot1 <- p1 + geom_line(aes(group = sample, colour = sample)) +
      theme(legend.position = "none")
    plot2 <- p2 + geom_line(aes(group = sample, colour = sample)) +
      theme(legend.position = "none")
    plot3 <- p3 + geom_line(aes(group = sample, colour = sample)) +
      theme(legend.position = "none")
    plot4 <- p4 + geom_line(aes(group = sample, colour = sample)) +
      theme(legend.position = "none")
  }
  # Making a grid display of the four plot objects above
  gridExtra::grid.arrange(plot1,plot2,plot3,plot4, ncol = 2,
               top = paste0("Simulated Spline Data (Sample Size = ", n,
                            ", Loops = ", loops, ", Weights = ", weights, ")"))
}
```

```{r WeightsSimsResults, results = FALSE, fig.width=7, fig.height=7}
Main.weights.plot <- Sims.Weight.plot(200, 10, "box", "samp_wt_quad")
```




## Sims: Use of surveyCV folds with SRS, clustered, or stratified samples


```{r FoldsSimsFunction}
# TODO: rename spline.df2 to be more clearly
# the sim dataset used for testing effect of folds (SRS vs clus vs strat)

# TODO: both here and for weights sims,
# separate the simulated-data code from the ggplot code, so that we can
# (1) store the raw data inside package instead of storing larger ggplot objects;
# (2) double-check things (like the warning about boxplot removing non-finite values)

clusVsrsVstrat.cross.cv.spline.plot <- function(n, loops) {
  
  # Use SRS samples and make SRS folds
  srssrsds <- data.frame(df = c(), MSE = c())
  # Use Cluster samples and make Cluster folds
  clusclusds <- data.frame(df = c(), MSE = c())
  # Use Cluster samples and make SRS folds
  clussrsds <- data.frame(df = c(), MSE = c())
  # Use Strat samples and make Strat folds
  stratstratds <- data.frame(df = c(), MSE = c())
  # Use Strat samples and make SRS folds
  stratsrsds <- data.frame(df = c(), MSE = c())
  
  modelsToFit <- c("Response~ns(Predictor, df=1)", "Response~ns(Predictor, df=2)",
                   "Response~ns(Predictor, df=3)", "Response~ns(Predictor, df=4)",
                   "Response~ns(Predictor, df=5)", "Response~ns(Predictor, df=6)")
  
  # Making as many simple random samples as we specify for 'loops'
  for (i in 1:loops) {
    set.seed(i)
    sim.srs <- dplyr::sample_n(spline.df2, n)
    # Using our SRS function on SRS samples
    srs.data <- cv.svy(sim.srs, modelsToFit,
                          nfolds = 5) %>% print()

    srssrsds2 <- data.frame(df = 1:6, MSE = srs.data[,1])
    srssrsds <- rbind(srssrsds, srssrsds2)
    # NOTE: removed srsclusds, since we don't need it for the paper
  }
  
  # Making as many cluster samples as we specify for 'loops'
  for (i in 1:loops) {
    set.seed(i)
    c <- unique(spline.df2[["Cluster"]])
    sim.clus <- spline.df2[spline.df2[["Cluster"]] %in% sample(c, n/10),]
    # Using our Cluster function on Cluster samples
    clus.data <- cv.svy(sim.clus, modelsToFit,
                               clusterID = "Cluster", nfolds = 5) %>% print()
    # Using our SRS function on Cluster samples
    srs.data <- cv.svy(sim.clus, modelsToFit,
                          nfolds = 5) %>% print()
    
    clusclusds2 <- data.frame(df = 1:6, MSE = clus.data[,1])
    clusclusds <- rbind(clusclusds, clusclusds2)
    clussrsds2 <- data.frame(df = 1:6, MSE = srs.data[,1])
    clussrsds <- rbind(clussrsds, clussrsds2)
  }
  
  # Making as many stratified samples as we specify for 'loops'
  for (i in 1:loops) {
    set.seed(i)
    s <- survey::stratsample(spline.df2$Stratum,
           c("1" = n/5, "2" = n/5, "3" = n/5, "4" = n/5, "5" = n/5))
    sim.strat <- spline.df2[s,]
    # Using our SRS function on Strat samples
    srs.data <- cv.svy(sim.strat, modelsToFit,
                       nfolds = 5) %>% print()
    
    # NOTE: Couldn't replicate plot_generation.R's strat sims exactly here,
    # since they use a separately-coded CV function...
    # and so, even when I tried running appendfolds() 4 times in this spot,
    # I still couldn't get the random seed to match up how it would look
    # after 6 iterations (in plot_gen...R) vs 2 iterations (here).
    # BUT it looks very similar and makes the same point -- just not identical.

    # Using our Strat function on Strat samples
    strat.data <- cv.svy(sim.strat, modelsToFit,
                         strataID = "Stratum", nfolds = 5) %>% print()
    
    stratstratds2 <- data.frame(df = 1:6, MSE = strat.data[,1])
    stratstratds <- rbind(stratstratds, stratstratds2)
    stratsrsds2 <- data.frame(df = 1:6, MSE = srs.data[,1])
    stratsrsds <- rbind(stratsrsds, stratsrsds2)
  }

  # Making the degrees of freedom variable a factor variable
  srssrsds$df <- as.factor(srssrsds$df)
  clusclusds$df <- as.factor(clusclusds$df)
  clussrsds$df <- as.factor(clussrsds$df)
  stratstratds$df <- as.factor(stratstratds$df)
  stratsrsds$df <- as.factor(stratsrsds$df)
  
  # Find the y-range of MSEs
  ymin <- min(min(srssrsds$MSE), min(clusclusds$MSE), min(clussrsds$MSE),
              min(stratstratds$MSE), min(stratsrsds$MSE))
  ymax <- max(max(srssrsds$MSE), max(clusclusds$MSE), max(clussrsds$MSE),
              max(stratstratds$MSE), max(stratsrsds$MSE))

  # Making ggplot objects for the MSEs
  plot1 <- ggplot2::ggplot(data = srssrsds, mapping = ggplot2::aes(x = df, y = MSE)) +
    ggplot2::geom_boxplot() +
    ggplot2::ggtitle("SRS folds with SRS sample") +
    scale_y_log10(limits = c(ymin, ymax))
  plot2 <- rectGrob(width = 0, height = 0) # empty spot here -- no corresponding sim
  plot3 <- ggplot2::ggplot(data = clusclusds, mapping = ggplot2::aes(x = df, y = MSE)) +
    ggplot2::geom_boxplot() +
    ggplot2::ggtitle("Cluster folds with Cluster sample") +
    scale_y_log10(limits = c(ymin, ymax))
  plot4 <- ggplot2::ggplot(data = clussrsds, mapping = ggplot2::aes(x = df, y = MSE)) +
    ggplot2::geom_boxplot() +
    ggplot2::ggtitle("SRS folds with Cluster sample") +
    scale_y_log10(limits = c(ymin, ymax))
  plot5 <- ggplot2::ggplot(data = stratstratds, mapping = ggplot2::aes(x = df, y = MSE)) +
    ggplot2::geom_boxplot() +
    ggplot2::ggtitle("Strat folds with Strat sample") +
    scale_y_log10(limits = c(ymin, ymax))
  plot6 <- ggplot2::ggplot(data = stratsrsds, mapping = ggplot2::aes(x = df, y = MSE)) +
    ggplot2::geom_boxplot() +
    ggplot2::ggtitle("SRS folds with Strat sample") +
    scale_y_log10(limits = c(ymin, ymax))

  
  # Making a grid display of the six plot objects above
  gridExtra::grid.arrange(plot1,plot2,plot4,plot3,plot6,plot5,
               ncol = 2,
               top = paste0("Simulated Spline Data (Sample Sizes = ", n,
                            ", Clusters = ", n/10,
                            " or Strata = 5, Loops = ", loops, ")"))
}
```

```{r FoldsSimsResults, results = FALSE, fig.width=7, fig.height=7}
sim.clusVsrsVstrat.sd.n100 <-
  clusVsrsVstrat.cross.cv.spline.plot(n=100, loops=100)
```




## Sims: Attempt to use dAIC with SRS, clustered, or stratified samples and compare to surveyCV

```{r, fig.width=7, fig.height=7}
n <- 100
loops <- 10

# Use SRS samples and make SRS folds
srssrsds <- data.frame(df = c(), MSE = c())
# Use SRS samples and use dAIC
srsdAICds <- data.frame(df = c(), dAIC = c())

modelsToFit <- c("Response~ns(Predictor, df=1)", "Response~ns(Predictor, df=2)",
                 "Response~ns(Predictor, df=3)", "Response~ns(Predictor, df=4)",
                 "Response~ns(Predictor, df=5)", "Response~ns(Predictor, df=6)")

# Making as many simple random samples as we specify for 'loops'
for (i in 1:loops) {
  set.seed(i)
  sim.srs <- dplyr::sample_n(spline.df2, n)
  # Using our SRS function on SRS samples
  srs.CV.data <- cv.svy(sim.srs, modelsToFit,
                        nfolds = 5) %>% print()
  srs.des <- svydesign(ids = 1:n, data = sim.srs, fpc = rep(n/1000, n))
  srs.dAIC.data <- sapply(1:6, function(ii) AIC(svyglm(modelsToFit[ii], srs.des))["AIC"])

  srssrsds2 <- data.frame(df = 1:6, MSE = srs.CV.data[,1])
  srssrsds <- rbind(srssrsds, srssrsds2)
  srsdAICds2 <- data.frame(df = 1:6, dAIC = srs.dAIC.data)
  srsdAICds <- rbind(srsdAICds, srsdAICds2)
}

srssrsds$df <- as.factor(srssrsds$df)
srsdAICds$df <- as.factor(srsdAICds$df)

plot1 <- ggplot(srssrsds, aes(x = df, y = MSE)) + geom_boxplot() +
    ggplot2::ggtitle("SRS folds with SRS sample")
plot2 <- ggplot(srsdAICds, aes(x = df, y = dAIC)) + geom_boxplot() +
    ggplot2::ggtitle("dAIC with SRS sample")

gridExtra::grid.arrange(plot1,plot2,
             ncol = 2,
             top = paste0("Simulated Spline Data (Sample Sizes = ", n,
                          ", Clusters = ", n/10,
                          " or Strata = 5, Loops = ", loops, ")"))
```

TODO: go back and add these dAIC sims to `clusVsrsVstrat.cross.cv.spline.plot()`,
showing them as a 3rd column of plots (one each for SRS, strat, and clus).
